---
title: "Intermediary Report - Capstone Project - word predictor"
author: "Longwei"
date: "4 juin 2016"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---


# Introduction and Objective

## Objective of the project

The goal of this project is to build a product (word predictor) which can predict the next word knowing the first words of a sentence. The final result of the project consists in an application (built with Shiny framework) using a model and algorithm based on machine learning and natural language processing tools (NLP). The model will be built on an data set consisting in extracts from US news, twitter and blogs.

## Overall approach

1 - Download the data, extract the zip file, make a summary of the file properties
2 - Load a randome sample of the data in R (as the data set is huge, this is necessary to save time and memory)
3 - Create a corpus, clean the data (Tokenise, remove profanity, ...)
4 - Exploratory data analysis
5 - Build the model
6 - Test the model
7 - Build a data product using the model to predict next words from the first words of a sentence
8 - Final report

## Objective of this document

This document is an intermediary report covering point 1 to 4 of previous section, focusing on exploratory analysis. We will also describe our general approach and goals for next steps (building the model, prediction algorithm and data product).

## Optimising performance and memory footprint

As this dataset is very large and we will have constraints to run the shiny app, as special focus should be taken to optimised the model for smaller memory footprint.

We found 3 alternatives of package for such process

- tm
- quanteda
- text2vec

## Notes
Explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. 

The motivation for this project is to: 

1. Demonstrate that you've downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app. 


# Load data

```{r configuration, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE, eval=TRUE}
# Clear objects in memory
rm(list = ls(all=TRUE))

# Load Libraries
library(data.table); library(R.utils); 
library(ggplot2); library(RColorBrewer); library(pander)

library(stringi); library(stringr)

library(quanteda); 
#library(NLP); library(openNLP)
#library(tokenizers); 

library("parallel"); library("foreach"); library("doParallel")
library(tm);library(RWeka);

library(text2vec); library(qdap); library(qdapRegex); library(SnowballC)
# Set a tracker data frame to record operation time
tracker <- data.frame(operation = "init", time = Sys.time())

# Garbage collection (memory)
gc()

# Multi core
#cl <- makeCluster(detectCores() - 1)
#registerDoParallel(cl, cores = detectCores() - 1)
```


## About HC Corpora 

The data is from a corpus called HC Corpora ([www.corpora.heliohost.org](www.corpora.heliohost.org)).
See the readme file at [http://www.corpora.heliohost.org/aboutcorpus.html](http://www.corpora.heliohost.org/aboutcorpus.html) for details on the corpora available. The files have been language filtered but may still contain some foreign text.

## Profanity Filtering

As required by the project assignement, we should filter profanity words. I propose to use a standard list found [online](https://www.google.com.hk/search?client=ubuntu&channel=fs&q=list+of+profanity+words&ie=utf-8&oe=utf-8&gfe_rd=cr&ei=ug1PV_mBGZSL8Qfr5pqQBA) for this task.

```{r getBadWords, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
# from "https://www.cs.cmu.edu/~biglou/resources/bad-words.txt
# http://www.bannedwordlist.com/lists/swearWords.txt"
badWordsUrl <- "../../data/swearWords.txt"
badWords <- read.csv(badWordsUrl,stringsAsFactors = FALSE)[,1]
tail(badWords)
```


## Overview the original data

Once the file is downloaded localy and unzip, we don't load it directly in R as the file are very large. We first use system level commands to characterise the data. This is basic information and do not take into account any data cleaning (the command wc -w counts words as token separated by spaces).

```{r overviewSystem, echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE, eval=TRUE, results='asis'}
blogsFile <-"../../data/final/en_US/en_US.blogs.txt"
newsFile <- "../../data/final/en_US/en_US.news.txt"
twitterFile <- "../../data/final/en_US/en_US.twitter.txt"

summaryDf <- data.frame(fileName = c(blogsFile,newsFile,twitterFile))
summaryDf$fileSize.Mo <- round(file.info(c(blogsFile,newsFile,twitterFile))$size / 1024^2)

# number of characters
summaryDf$nbChars <- c(
        strsplit(system(paste("wc -m ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -m ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -m ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# number of lines
summaryDf$nbLines <- c(
        strsplit(system(paste("wc -l ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -l ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -l ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# max line lenght
summaryDf$MaxLineLength <- c(
        strsplit(system(paste("wc -L ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -L ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -L ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# number of words, knowing wc is only doing a basic tokenisation
summaryDf$nbWords <- c(
        strsplit(system(paste("wc -w ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -w ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -w ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# convert to numeric for later calculations
summaryDf$nbLines <- as.numeric(summaryDf$nbLines)

pandoc.table(summaryDf, style = "rmarkdown", split.table = Inf)

tracker <- rbind(tracker, data.frame(operation = "Summarize Data (with wc)", time = Sys.time()))
```




## Read the files by Sampling

As the data set is very large, we reach the limit of RAM and computing very quickly, we decide to subset a random sample of the initial data to make the exploraroty analysis and build the model. We start with 1000 lines to developp our full process and we will later iterate by increasing `sampleSize` to improve accuracy.


```{r loadData, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
set.seed(1234)
sampleSize <- 0.005

con <- file(blogsFile, "r") 
blogsSample <- sample(
        readLines(con, -1, skipNul = TRUE),
        round(sampleSize*summaryDf$nbLines[1]))
close(con)
con <- file(newsFile, "r")
newsSample <- sample(
        readLines(con, -1, skipNul = TRUE), 
        round(sampleSize*summaryDf$nbLines[2]))
close(con) 
con <- file(twitterFile , "r")
twitterSample <- sample(
        readLines(con, -1, skipNul = TRUE),
        round(sampleSize*summaryDf$nbLines[3]))
close(con)

tracker <- rbind(tracker, data.frame(operation = paste("Load Data Sample :", sampleSize), time = Sys.time()))
rm(summaryDf)
```

# Overview of the sampled data

```{r overViewOfData, message=FALSE, warning=FALSE, eval=TRUE, cache=FALSE, echo=TRUE}
tail(blogsSample,3)
tail(newsSample,3)
tail(twitterSample,5)
```


## Summary of the data with summary(corpus) from quanteda

```{r summaryDatSetv2, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
enSample <- corpus(
        c(
                blogsSample = paste(blogsSample, collapse = " "),
                newsSample = paste(newsSample, collapse = " "),
                twitterSample = paste(twitterSample, collapse = "")
                )
        )
tracker <- rbind(tracker, data.frame(operation = "Create Corpus from Sample data", time = Sys.time()))
```

```{r summaryDatSetv3, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE, results='asis'}
summaryEnSample <- summary(enSample)

pandoc.table(summaryEnSample, style = "rmarkdown", split.table = Inf)

tracker <- rbind(tracker, data.frame(operation = "Summarize Sample data - quanteda", time = Sys.time()))
```

## Corpus overview

```{r  graphCorpus, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
## Graphical summary
g <- ggplot(data = summaryEnSample)
g <- g + geom_point(aes(x = Tokens, y = Types, size = Sentences, col = Text))
g <- g + scale_size_area(max_size = 15)
g
```


# Text with text2vec

## What is text2vec ?
http://dsnotes.com/articles/text2vec-0-3

## Vocabulary based vectorization

### Steming and Tokenizer function 

```{r text2vecTokenizer, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
stem_tokenizer <- function(v, tokenizer = word_tokenizer) {
  v %>%
    tokenizer %>%
    # poerter stemmer
    lapply(wordStem, 'en')
    # Remove numbers ?
    #%>% removeNumbers
}
sw <- c(stopwords("en"), badWords)
```

### Stop words and bad words

```{r text2vecMakeTokens, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}

tokens <- c(paste(blogsSample, collapse = " "),
            paste(newsSample, collapse = " "),
            paste(twitterSample, collapse = "")
            ) %>%
        tolower %>%
        stem_tokenizer
names(tokens) <- c("blogsSample", "newsSample", "twitter")

## Full stopword and badword list
## swF <- c(stopwords("en"), badWords)
## Short stopword and badword list
swC <- c(
        c("i", "me", "my", "myself", "we", "our", "ours", "ourselves", "you", "your", "yours"),
        badWords) %>%
# here we stem stopwords, because stop-words filtering would be performed after tokenization!
        wordStem('en')
```

## Plotting Functions 

```{r plotFunctions,  message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
makeTopFeaturesDfm <- function(aDfm,maxTopic){
        top <- as.data.frame(topfeatures(aDfm, maxTopic))
        names(top) <- c("freq"); top$feature <- rownames(top)
        rownames(top) <- NULL
        top <- dplyr::select(top, feature, freq)
        top
}

plotFeatures <- function(featureDf) {
        g <- ggplot(data = featureDf)
        g <- g + geom_bar(aes(x = reorder(term, count), y = count), stat = "identity")
        g <- g +  coord_flip()
        g
}
```


## UniGrams

```{r text2vecUniGrams, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
## Create the iterator on the tokens

it <- itoken(tokens)
vocab <- create_vocabulary(it, ngram = c(ngram_min = 2L, ngram_max = 2L), stopwords = swC)



N_WORKERS <- detectCores() - 1
library(doParallel)
library(microbenchmark)
registerDoParallel(N_WORKERS)
# "jobs" is a list of itoken iterators!
N_SPLITS <- detectCores() - 1
jobs <- tokens %>% 
  split_into(N_SPLITS) %>% 
  lapply(itoken)
# performance comparison between serial and parallel versions
# vocabulary creation
microbenchmark(
  vocab_serial <- create_vocabulary(itoken(tokens), ngram = c(ngram_min = 3L, ngram_max = 3L), stopwords = swC), 
  vocab_parallel <- create_vocabulary(jobs, ngram = c(ngram_min = 3L, ngram_max = 3L), stopwords = swC), 
  times = 1
)


#So before corpus / dtm / tcm construction we need to reinitialise it. Here we create dtm directly:
it <- itoken(tokens)
v_vectorizer <- vocab_vectorizer(vocab_parallel)
dtm <- create_dtm(it, v_vectorizer)
# check  that dtm keep documents names/ids as rownames
head(rownames(dtm))

words <- colSums(dtm)
words <- words[order(words, decreasing = T)][1:100]
wordcloud::wordcloud(names(words), words, colors = brewer.pal(6, "Dark2"), scale = c(8, .5))



#dt1 <- data.table(term = vocab1$vocab$terms, count = vocab1$vocab$terms_count)
#dt1 <- dt1[order(-count)]
tracker <- rbind(tracker, data.frame(operation = "Create Vocabulary Unigram - text2vec", time = Sys.time()))
```


```{r wordCloudUni, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
wordcloud::wordcloud(dt1[1:100,term], dt1[1:100,count], colors = brewer.pal(6, "Dark2"), scale = c(8, .5))
```

```{r featurePlotUniGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
plotFeatures(dt1[order(-count)][1:20])

save(vocab1, file = '../../data/Rda/vocab1.Rda')
rm(vocab1,dt1)
gc()
```

## BiGrams

```{r text2vecBiGrams, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
it <- itoken(list(
                paste(blogsSample, collapse = " "),
                paste(newsSample, collapse = " "),
                paste(twitterSample, collapse = "")
                ), 
             preprocess_function = preprocess , 
             tokenizer = word_tokenizer)
vocab2 <- create_vocabulary(it, stopwords = sw, ngram = c(ngram_min = 2L, ngram_max = 2L))
dt2 <- data.table(term = vocab2$vocab$terms, count = vocab2$vocab$terms_count)
dt2 <- dt2[order(-count)]
tracker <- rbind(tracker, data.frame(operation = "Create Vocabulary Bigram - text2vec", time = Sys.time()))

```


```{r wordCloudBiGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
wordcloud::wordcloud(dt2[1:100,term], dt2[1:100,count], colors = brewer.pal(6, "Dark2"), scale = c(8, .5))
```

```{r featurePlotBiGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
plotFeatures(dt2[order(-count)][1:20])

save(vocab2, file = '../../data/Rda/vocab2.Rda')
rm(vocab2,dt2)
gc()
```


## TriGrams

```{r text2vecTriGrams, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
it <- itoken(list(
                paste(blogsSample, collapse = " "),
                paste(newsSample, collapse = " "),
                paste(twitterSample, collapse = "")
                ), 
             preprocess_function = preprocess , 
             tokenizer = word_tokenizer)
vocab3 <- create_vocabulary(it, stopwords = sw, ngram = c(ngram_min = 3L, ngram_max = 3L))
dt3 <- data.table(term = vocab3$vocab$terms, count = vocab3$vocab$terms_count)
dt3 <- dt3[order(-count)]
tracker <- rbind(tracker, data.frame(operation = "Create Vocabulary Trigram - text2vec", time = Sys.time()))

```


```{r wordCloudTriGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
wordcloud::wordcloud(dt3[1:100,term], dt3[1:100,count], colors = brewer.pal(6, "Dark2"), scale = c(8, .5))
```

```{r featurePlotTriGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
plotFeatures(dt3[order(-count)][1:20])

save(vocab3, file = '../../data/Rda/vocab3.Rda')
rm(vocab3,dt3)
gc()
```




# Exploratory Analysis

## Exploratory analysis
The first step in building a predictive model for text is understanding the distribution and relationship between the words, tokens, and phrases in the text. The goal of this task is to understand the basic relationships you observe in the data and prepare to build your first linguistic models.

Exploratory analysis - perform a thorough exploratory analysis of the data, understanding the distribution of words and relationship between the words in the corpora.
Understand frequencies of words and word pairs - build figures and tables to understand variation in the frequencies of words and word pairs in the data.

Some words are more frequent than others - what are the distributions of word frequencies?
What are the frequencies of 2-grams and 3-grams in the dataset?
How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 90%?
How do you evaluate how many of the words come from foreign languages?
Can you think of a way to increase the coverage -- identifying words that may not be in the corpora or using a smaller number of words in the dictionary to cover the same number of phrases?










## Processing time

```{r processingDuration, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE, results='asis'}
tracker$op.duration <- round(c(0,(tracker[2:nrow(tracker),]$time - tracker[1:(nrow(tracker)-1),]$time)),1)
tracker$total.duration <- round(tracker$time - tracker[1,]$time,1)
pandoc.table(tracker, style = "rmarkdown", split.table = Inf)
```


# Approacha for prediction


# Conclusion
```{r}
stopCluster(cl)
```


# References

Review Criteria :

- Does the link lead to an HTML page describing the exploratory analysis of the training data set?
- Has the data scientist done basic summaries of the three files? Word counts, line counts and basic data tables?
- Has the data scientist made basic plots, such as histograms to illustrate features of the data?
- Was the report written in a brief, concise style, in a way that a non-data scientist manager could appreciate?