---
title: "Intermediary Report - Capstone Project - word predictor"
author: "Longwei"
date: "4 juin 2016"
output: 
  html_document: 
    number_sections: yes
    toc: yes
---


# Introduction and Objective

## Objective of the project

The goal of this project is to build a product (word predictor) which can predict the next word knowing the first words of a sentence. The final result of the project consists in an application (built with Shiny framework) using a model and algorithm based on machine learning and natural language processing tools (NLP). The model will be built on an data set consisting in extracts from US news, twitter and blogs.

## Overall approach

1 - Download the data, extract the zip file, make a summary of the file properties.
2 - Load a random sample of the data in R (as the data set is huge, this is necessary to save time and memory).
3 - Clean the data, tokenise (extract words).
4 - Compute frequency for n-grams (sequence of 1 to n words : 1 to 3 in this report.
5 - Exploratory data analysis.
6 - Build the model.
7 - Test the model.
8 - Build a data product using the model to predict next words from the first words of a sentence.
9 - Final report.

## Objective of this document

This document is an intermediary report covering point 1 to 5 of previous section, focusing on exploratory analysis. We will also describe our general approach and goals for next steps (building the model, prediction algorithm and data product).

## Optimising performance and memory footprint

As this dataset is very large and we will have constraints to run the shiny app, as special focus should be taken to optimised the model for smaller memory footprint.

We found 3 alternatives of package for text mining :

- tm, provide plenty of functions but its documentation is not straightforward and we faced memory issues for n-grams extractions
- quanteda, a more recent package with better performance. This is the package we will use for basic data about the corpus of text
- text2vec, the latest package in the text mining familly. Its documentation is quite limited but the performances are much better than the previous ones as it is coded in C and optimised for parallel processing. This is the package we use for computing n-grams.

## Notes
Explain only the major features of the data you have identified and briefly summarize your plans for creating the prediction algorithm and Shiny app in a way that would be understandable to a non-data scientist manager. 

The motivation for this project is to: 

1. Demonstrate that you've downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app. 


# Load data

```{r configuration, echo=FALSE, warning=FALSE, message=FALSE, cache=FALSE, eval=TRUE}
# Clear objects in memory
rm(list = ls(all=TRUE))

# Load Libraries
library(data.table); library(R.utils); library(dplyr)
library(ggplot2); library(RColorBrewer); library(pander)
# For Parallele Computing
library("parallel"); library("foreach"); library("doParallel")

library(stringi); library(stringr); library(quanteda); 
library(tm);library(RWeka);
library(text2vec); library(qdap); library(SnowballC)
# Set a tracker data frame to record operation time
tracker <- data.frame(operation = "init", time = Sys.time())

# Garbage collection (memory)
gc()
# Multi core
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores = detectCores() - 1)
```


## About HC Corpora 

The data is from a corpus called HC Corpora ([www.corpora.heliohost.org](www.corpora.heliohost.org)).
See the readme file at [http://www.corpora.heliohost.org/aboutcorpus.html](http://www.corpora.heliohost.org/aboutcorpus.html) for details on the corpora available. The files have been language filtered but may still contain some foreign text.

## Profanity Filtering

As required by the project assignement, we should filter profanity words. We propose to use a standard list found online for this task.

```{r getBadWords, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
# http://www.bannedwordlist.com/lists/swearWords.txt"
# another more detailled here : # from "https://www.cs.cmu.edu/~biglou/resources/bad-words.txt
badWordsUrl <- "../../data/swearWords.txt"
badWords <- read.csv(badWordsUrl,stringsAsFactors = FALSE)[,1]
tail(badWords)
```


## Overview the original data

Once the assignment text data file is downloaded localy and unziped, we don't load it directly in R as the file are very large. We first use system level commands to characterise the data. This is basic information and do not take into account any data cleaning (the command wc -w counts words as token separated by spaces).

```{r overviewSystem, echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE, eval=TRUE, results='asis'}
blogsFile <-"../../data/final/en_US/en_US.blogs.txt"
newsFile <- "../../data/final/en_US/en_US.news.txt"
twitterFile <- "../../data/final/en_US/en_US.twitter.txt"

summaryDf <- data.frame(fileName = c(blogsFile,newsFile,twitterFile))
summaryDf$fileSize.Mo <- round(file.info(c(blogsFile,newsFile,twitterFile))$size / 1024^2)

# number of characters
summaryDf$nbChars <- c(
        strsplit(system(paste("wc -m ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -m ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -m ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# number of lines
summaryDf$nbLines <- c(
        strsplit(system(paste("wc -l ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -l ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -l ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# max line lenght
summaryDf$MaxLineLength <- c(
        strsplit(system(paste("wc -L ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -L ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -L ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# number of words, knowing wc is only doing a basic tokenisation
summaryDf$nbWords <- c(
        strsplit(system(paste("wc -w ", blogsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -w ", newsFile), intern = TRUE), split = " ")[[1]][1],
        strsplit(system(paste("wc -w ", twitterFile), intern = TRUE), split = " ")[[1]][1]
        )
# convert to numeric for later calculations
summaryDf$nbLines <- as.numeric(summaryDf$nbLines)

pandoc.table(summaryDf, style = "rmarkdown", split.table = Inf)

tracker <- rbind(tracker, data.frame(operation = "Summarize Data (with wc)", time = Sys.time()))
```




## Read the files by Sampling

As the data set is very large, we reach the limit of RAM and computing very quickly, we decide to subset a random sample of the initial data to make the exploraroty analysis and build the model. We start with few % of each file to developp our full process and we will later iterate by increasing `sampleSize` to improve accuracy.


```{r loadData, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
set.seed(1234)
sampleSize <- c(0.01,0.01,0.01)

con <- file(blogsFile, "r") 
blogsSample <- sample(
        readLines(con, -1, skipNul = TRUE),
        round(sampleSize[1]*summaryDf$nbLines[1]))
close(con)
con <- file(newsFile, "r")
newsSample <- sample(
        readLines(con, -1, skipNul = TRUE), 
        round(sampleSize[2]*summaryDf$nbLines[2]))
close(con) 
con <- file(twitterFile , "r")
twitterSample <- sample(
        readLines(con, -1, skipNul = TRUE),
        round(sampleSize[3]*summaryDf$nbLines[3]))
close(con)

tracker <- rbind(tracker, data.frame(operation = paste("Load Data Sample :", sampleSize), time = Sys.time()))
rm(summaryDf)
```

# Overview of the sampled data

```{r overViewOfData, message=FALSE, warning=FALSE, eval=TRUE, cache=FALSE, echo=TRUE}
tail(blogsSample,3)
tail(newsSample,3)
tail(twitterSample,5)
```


## Summary of the Sampled data with summary(corpus) from quanteda

We use quanteda summary function on a corpus to analyse the key dimensions of our data Sample.

```{r summaryDatSetv2, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
enSample <- corpus(
        c(
                blogsSample = paste(blogsSample, collapse = " "),
                newsSample = paste(newsSample, collapse = " "),
                twitterSample = paste(twitterSample, collapse = "")
                )
        )
tracker <- rbind(tracker, data.frame(operation = "Create Corpus from Sample data", time = Sys.time()))
```

```{r summaryDatSetv3, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE, results='asis'}
summaryEnSample <- summary(enSample)

pandoc.table(summaryEnSample, style = "rmarkdown", split.table = Inf)

tracker <- rbind(tracker, data.frame(operation = "Summarize Sample data - quanteda", time = Sys.time()))
```

The following plot helps to visualise the different text. Twitter texts have much more different Types (different words) than the News of Blogs data. This seems logical as using twitter, users are force to condense words and use hachtags, etc... Twitter text data base lines are much shorter which explains there are less tokens in this dataset than in the news or blogs one.

```{r  graphCorpus, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
## Graphical summary
g <- ggplot(data = summaryEnSample)
g <- g + geom_point(aes(x = Tokens, y = Types, size = Sentences, col = Text))
g <- g + scale_size_area(max_size = 15)
g <- g + ggtitle("Key parameters of the thre text corpus of the english document database")
g
```


# Tokent extraction of english database with text2vec

## What is text2vec ?
From https://github.com/dselivanov/text2vec

> text2vec is a package that provides an efficient framework with a concise API for text analysis and natural language processing (NLP) in R. It is inspired by gensim, an excellent Python library for NLP. [...] The core of this package is carefully written in C++, which means text2vec is fast and memory friendly. 

## Vocabulary based vectorization

In order to do an exploratory analysis of the corpus and then build our prediction model, the first step is to vectorise the corpus. It means first to extract the different tokens and to compute their frequency in the 3 types of ressources.

### Steming, pre-processing and tokenizer function 

First, we build a function to extract tokens for the text using a tokenizer and some preprocessing. For this intermediary report, we will use a limited amount of pre-processing :

- convert text to lower case
- remove english stopwords
- remove bad words (profanity filtering)

Some other filtering to be considered later to build the model :

- Stemming ([wikipedia](https://en.wikipedia.org/wiki/Stemming) : A stemmer for English, for example, should identify the string "cats" (and possibly "catlike", "catty" etc.) as based on the root "cat", and "stems", "stemmer", "stemming", "stemmed" as based on "stem". A stemming algorithm reduces the words "fishing", "fished", and "fisher" to the root word, "fish".)
- url filtering (remove all uri)
- twitter filtering (how to deal with hashtags, etc..)
- remove numbers ?


```{r text2vecTokenizer, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
stem_tokenizer <- function(v, tokenizer = word_tokenizer) {
  v %>%
    tokenizer 
    # poerter stemmer
    #  %>% lapply(wordStem, 'en')
    # Remove numbers ?
    # %>% removeNumbers
}
sw <- c(stopwords("en"), badWords)
```

We apply the stem_tokenizer function to our data set.

```{r text2vecMakeTokens, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
tokens <- c(paste(blogsSample, collapse = " "),
            paste(newsSample, collapse = " "),
            paste(twitterSample, collapse = "")
            ) %>%
        tolower %>%
        stem_tokenizer
names(tokens) <- c("blogsSample", "newsSample", "twitter")


```

### Stopwords, bad words filtering

Filtering the english stopwords and bad words will be required in the next steps. We prepare the merge list with the following code.

```{r text2vecBadWords, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
## stopwords are words like : c("i", "me", "my", "myself", "we", "our", "ours", "ourselves", "you", "your", "yours")
swC <- c(stopwords("en"), 
        badWords) 
# here if we use stemming; we need also to stem stopwords, because stop-words filtering would be performed after tokenization!
# %>% wordStem('en')
```


## UniGrams

```{r text2vecUniGrams, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
## Create the iterator on the tokens
it <- itoken(tokens)
## Generate the vocabulary using the iterator, 1-gram badwords / stopwords list.
vocab1 <- create_vocabulary(it, ngram = c(ngram_min = 1L, ngram_max = 1L), stopwords = swC)

## We need to reinitialise iterator : it
## Here we create dtm directly:
it <- itoken(tokens)
v_vectorizer <- vocab_vectorizer(vocab1)
dtm1 <- create_dtm(it, v_vectorizer)
# check  that dtm keep documents names/ids as rownames
rownames(dtm1)

## Save dtm1 R object
save(dtm1, file = '../../data/Rda/dtm1.Rda')

## Sum freq for each document / feature
wordsFreq <- colSums(dtm1)

## Extract the top 100 most frequent concepts
top1 <- wordsFreq[order(wordsFreq, decreasing = T)][1:100]
top1 <- data.frame(features = names(top1),
                   freq = top1, row.names = NULL)
top1 <- transform( 
        top1,
        features = reorder(features,freq))
## Order factor by freq
top1$features <-factor(top1$features, levels=top1[order(top1$freq), "features"])


tracker <- rbind(tracker, data.frame(operation = "Create Vocabulary / DTM Unigram - text2vec", time = Sys.time()))
```

## Bi Grams

```{r text2vecBiGrams, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
## Create the iterator on the tokens
it <- itoken(tokens)
## Generate the vocabulary using the iterator, 1-gram badwords / stopwords list.
vocab2 <- create_vocabulary(it, ngram = c(ngram_min = 2L, ngram_max = 2L), stopwords = swC)

## We need to reinitialise iterator : it
## Here we create dtm directly:
it <- itoken(tokens)
v_vectorizer <- vocab_vectorizer(vocab2)
dtm2 <- create_dtm(it, v_vectorizer)
# check  that dtm keep documents names/ids as rownames
rownames(dtm2)
object.size(dtm2) / 1024^2

## Save dtm1 R object
save(dtm2, file = '../../data/Rda/dtm2.Rda')

## Sum freq for each document / feature
wordsFreq <- colSums(dtm2)

## Extract the top 100 most frequent concepts
top2 <- wordsFreq[order(wordsFreq, decreasing = T)][1:100]
top2 <- data.frame(features = names(top2),
                   freq = top2, row.names = NULL)
## Order factor by freq
top2$features <-factor(top2$features, levels=top2[order(top2$freq), "features"])


tracker <- rbind(tracker, data.frame(operation = "Create Vocabulary Bigram - text2vec", time = Sys.time()))
```

## Tri Grams

```{r text2vecTriGrams, message=FALSE, warning=FALSE, cache=FALSE, echo = TRUE, eval=TRUE}
## Create the iterator on the tokens
it <- itoken(tokens)
## Generate the vocabulary using the iterator, 1-gram badwords / stopwords list.
vocab3 <- create_vocabulary(it, ngram = c(ngram_min = 3L, ngram_max = 3L), stopwords = swC)

## We need to reinitialise iterator : it
## Here we create dtm directly:
it <- itoken(tokens)
v_vectorizer <- vocab_vectorizer(vocab3)
dtm3 <- create_dtm(it, v_vectorizer)
# check  that dtm keep documents names/ids as rownames
rownames(dtm3)
object.size(dtm3) / 1024^2

## Save dtm1 R object
save(dtm3, file = '../../data/Rda/dtm2.Rda')

## Sum freq for each document / feature
wordsFreq <- colSums(dtm3)

## Extract the top 100 most frequent concepts
top3 <- wordsFreq[order(wordsFreq, decreasing = T)][1:100]
top3 <- data.frame(features = names(top3),
                   freq = top3, row.names = NULL)
## Order factor by freq
top3$features <-factor(top3$features, levels=top3[order(top3$freq), "features"])



tracker <- rbind(tracker, data.frame(operation = "Create Vocabulary Trigram - text2vec", time = Sys.time()))

```

# Exploratory Analysis

## Principles

## Exploratory analysis
The first step in building a predictive model for text is understanding the distribution and relationship between the words, tokens, and phrases in the text. The goal of this task is to understand the basic relationships you observe in the data and prepare to build your first linguistic models.

Exploratory analysis - perform a thorough exploratory analysis of the data, understanding the distribution of words and relationship between the words in the corpora.
Understand frequencies of words and word pairs - build figures and tables to understand variation in the frequencies of words and word pairs in the data.

Some words are more frequent than others - what are the distributions of word frequencies?
What are the frequencies of 2-grams and 3-grams in the dataset?
How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 90%?
How do you evaluate how many of the words come from foreign languages?
Can you think of a way to increase the coverage -- identifying words that may not be in the corpora or using a smaller number of words in the dictionary to cover the same number of phrases?

## Plotting Functions 

```{r plotFunctions,  message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE}
plotFeatures <- function(featureDf) {
        g <- ggplot(data = featureDf)
        g <- g + geom_bar(aes(x = features, y = freq), stat = "identity")
        g <- g +  coord_flip()
        g
}
```


## Exploring Uni-grams

```{r wordCloudUni, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
set.seed(1234)
wordcloud::wordcloud(top1$features, top1$freq, colors = brewer.pal(6, "Dark2"), scale = c(8, .1))
```

```{r featurePlotUniGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
plotFeatures(top1[1:30,])
```

## Exploring BiGrams


```{r wordCloudBiGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
set.seed(1234)
wordcloud::wordcloud(top2$features, top2$freq, colors = brewer.pal(6, "Dark2"), scale = c(8, .1))
```

```{r featurePlotBiGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
plotFeatures(top2[1:30,])
```


## Exploring TriGrams


```{r wordCloudTriGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
set.seed(1234)
wordcloud::wordcloud(top3$features, top2$freq, colors = brewer.pal(6, "Dark2"), scale = c(8, .1))
```

```{r featurePlotTriGram, message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE, eval=TRUE}
plotFeatures(top3[1:30,])
```




# Appendix

## Processing time

```{r processingDuration, message=FALSE, warning=FALSE, echo=TRUE, eval=TRUE, cache=FALSE, results='asis'}
tracker$op.duration <- round(c(0,(tracker[2:nrow(tracker),]$time - tracker[1:(nrow(tracker)-1),]$time)),1)
tracker$total.duration <- round(tracker$time - tracker[1,]$time,1)
pandoc.table(tracker, style = "rmarkdown", split.table = Inf)
```


# Approacha for prediction


# Conclusion
```{r}
stopCluster(cl)
```


# References

Review Criteria :

- Does the link lead to an HTML page describing the exploratory analysis of the training data set?
- Has the data scientist done basic summaries of the three files? Word counts, line counts and basic data tables?
- Has the data scientist made basic plots, such as histograms to illustrate features of the data?
- Was the report written in a brief, concise style, in a way that a non-data scientist manager could appreciate?